---
title: |
  | Applications of R for Finance
  | 
author: |
  | 
  | 
  | 
date: "Date:`r Sys.Date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

\newpage
\tableofcontents
\listoftables
\newpage

Part I 

Load R packages

Load the packages required. Consider using `dplyr` , `lubridate` , `knitr`

```{r loadLibs, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)
library(lubridate)

```

Loading Data

Load `compustat_sec_2020_2023.csv` into a data frame.

```{r loadData}
raw_data <- read.csv("compustat_sec_2020_2023.csv")
```

Calculating Returns

Calculate simple daily returns for each stock.
**Ensure there are no `NA` values or `Inf` values in your returns column**

```{r calculateReturns}
data <- raw_data %>% 
  # Sorting and arranging the row called 'tic' based on the 'datadate' column
  arrange(tic, datadate) %>%
  # Grouping data based on variable 'tic'
  group_by(tic) %>% 
  # Enables the creation of a new variable called 'returns', which is the daily
  # return as a percentage change. It divides the 'prccd' column by the lagged 
  # 'prccd' (the previous day's closing price) and subtracts 1 from the result.
  mutate(returns = prccd / lag(prccd) -1) %>%
  # Modify the 'datadate' column to make it the correct format
  mutate(datadate = as.Date(datadate, format = "%Y-%m-%d")) %>% 
  # Remove grouping
  ungroup %>% 
  # To omit the 'NA' values
  na.omit()

# To make sure there is no 'Inf' values
data <- data[is.finite(data$returns), ]
# Check the first few rows of the data frame
head(data)

```

Create a new data frame containing the maximum daily return for each month, for
each stock. For each stock in the data frame, each row should represent the
maximum daily return in a given month.

(i.e., select the row with the maximum return in each month, for each stock)

**NOTES**:
- Keep all the columns of the original data frame

- Keep only one observation per month, per stock (in case of two same returns)
  - Remove the duplicates / keep only the first observation

```{r monthlyMaxReturns}
# Filter data into a new data frame and create a new column to indicate the month
data <- data %>% 
  mutate(month = floor_date(datadate, unit = "month"))

# Calculate Monthly mean returns into a new data frame
data_monthly <- data %>% 
  # Group your data by the "tic" and "month" columns
  group_by(tic, month) %>% 
  # Add a new column called 'max_returns' to store the max value of returns
  mutate(max_returns = max(returns)) %>% 
  # Return unique combinations of values for 'tic' and 'month' columns while 
  # preserving all other columns in the original data frame.
  distinct(tic, month, .keep_all = TRUE) %>%  
  # Remove grouping
  ungroup()

# Omit the NA values
data_monthly <- na.omit(data_monthly)
# Omit the infinite values
data_monthly[is.finite(data_monthly$returns), ]

# Check the first few rows
head(data_monthly)

```

Part II 

Perform a decile group analysis on a specific
month within the data set. Create a new data frame from the
data frame created. Categorise the maximum daily
returns from that month into decile groups, labelled 10%, 20%, ..., 90%, 100% .

Identify a stock in a certain decile based on
some criteria. The decile and criteria will be given in each section below.

**NOTES**
- For the quantile algorithm, use `type = 5`
- Use functions to create `probs` and `labels`

A

Consider only the month of March 2023, and create decile groups.
```{r decilesA}
# Create a new data frame 'dataA' based on 'data'
dataA <- data_monthly %>% 
  # Add a new column 'month' containing the start date of the month for each 
  # corresponding 'datadate' value
  mutate(month = floor_date(datadate, unit = "month")) %>% 
  # Filter rows where the 'month' column is equal to '2023-03-01', for March 2023
  filter(month == "2023-03-01") %>% 
  # Filter rows without zero returns
  filter(returns != 0) %>% 
  # Filter rows with finite returns value
  filter(is.finite(returns))

# Remove the NA values
dataA <- na.omit(dataA)
  
# Create a new variable that categorizes the values in 'returns' column into percentiles
dataA$percentiles <- cut(dataA$returns,
                         breaks = quantile(dataA$returns,
                                           probs = seq(0, 1, by = 0.1),
                                           type = 5,
                                           na.rm = TRUE),
                         include.lowest = TRUE,
                         labels = paste0(seq(10, 100, by = 10),"%"))
# Check for the first few rows
head(dataA)
```

Choose the stock matching the following criteria:
- lowest return in the 100% decile group
  
Display the stock in a pretty table with 8 columns showing this information:
- GVKEY, iid, date, company name, trading volume, closing price, return, decile

```{r selectA}
# Create a new data frame
lowest_dataA <- dataA %>% 
  # Filter the rows for 100% percentiles
  filter(percentiles == "100%") %>% 
  # Arrange the data frame based on 'returns' and then extract the first row. 
  arrange(returns) %>% 
  .[1, ]

lowest_dataA <- lowest_dataA %>% 
  # Select the following 8 colums
  select(tic, GVKEY, iid, datadate, conm, cshtrd, prccd, returns, percentiles)


# Create a table to display 'lowest_dataA' in a more readable format.
kable(lowest_dataA, caption = "Lowest return in the 10th Decile for March 2023")
```



B

Consider only the month of February 2023, and create decile groups.
```{r decilesB}
# Create a new data frame 'dataB' based on 'data'
dataB <- data_monthly %>% 
  # Add a new column 'month' containing the start date of the month for each 
  # corresponding 'datadate' value
  mutate(month <- format(datadate, "%m")) %>% 
  # Filter rows where the 'month' column is equal to '2023-02-01', for February 2023
  filter(month == "2023-02-01") %>% 
  # Filter rows without zero returns
  filter(returns != 0) %>% 
  # Filter rows with finite returns value
  filter(is.finite(returns))

# Filter rows with NA value
dataB <- na.omit(dataB)

# Create a new variable called 'percentiles' that categorizes the values in 
# 'returns' column into percentiles
dataB$percentiles <- cut(dataB$returns,
                         breaks = quantile(dataB$returns,
                                           probs = seq(0, 1, by = 0.1),
                                           type = 5,
                                           na.rm = TRUE),
                         include.lowest = TRUE,
                         labels = paste0(seq(10, 100, by = 10),"%"))
# Check the front few rows
head(dataB)
```

Choose the stock matching the following criteria:
- lowest return in the 80% decile group
  
Display the stock in a pretty table with 8 columns showing this information:
- GVKEY, iid, date, company name, trading volume, closing price, return, decile

```{r selectB}
# Create a new data frame
lowest_dataB <- dataB %>% 
  # Filter the rows for 80% percentiles
  filter(percentiles == "80%") %>% 
  # Arrange the data frame based on 'returns'.
  arrange(returns) %>%
  # Select only the following 8 columns
  select(tic,GVKEY, iid, datadate, conm, cshtrd, prccd, returns, percentiles) %>% 
  # Display the first row of' data frame'lowest_dataB'
  head(n = 1)

# Create a table to display 'lowest_dataB' in a more readable format.
kable(lowest_dataB, caption = "Lowest return in the 8th Decile for February 2023")


```

C

Consider only the month of January 2023, and create decile groups.
```{r decilesC}
# Filter data into a new data frame and create a new column to indicate the month
dataC <- data_monthly %>% 
  mutate(month <- format(datadate, "%m")) %>% 
  filter(month == "2023-01-01") %>% 
  filter(returns != 0) %>% 
  filter(is.finite(returns))

# Remove NA values
dataC <- na.omit(dataC)

# Create a new column to indicate the decile group
dataC$percentiles <- cut(dataC$returns,
                         breaks = quantile(dataB$returns,
                                           probs = seq(0, 1, by = 0.1),
                                           type = 5,
                                           na.rm = TRUE),
                         include.lowest = TRUE,
                         labels = paste0(seq(10, 100, by = 10),"%"))

head(dataC)
```

Choose the stock matching the following criteria:
- highest return in the 20% decile group
  
Display the stock in a pretty table with 8 columns showing this information:
- GVKEY, iid, date, company name, trading volume, closing price, return, decile

```{r selectC}
# Select the stock with the highest return in the 20% decile group
highest_dataC <- dataC %>% 
  filter(percentiles == "20%") %>% 
  arrange(desc(returns)) %>%
  select(tic, GVKEY, iid, datadate, conm, cshtrd, prccd, returns, percentiles) %>% 
  # Display the first row of' data frame'highest_dataD'
  head(n = 1)
highest_dataC
# Display the stock in a pretty table
kable(highest_dataC, caption = "Highest return in the 2nd Decile for January 2023")
```

D

Consider only the month of December 2022, and create decile groups.
```{r decilesD}
# Create a new data frame 'dataD'.
dataD <- data_monthly %>% 
  # Create a new variable called 'month' based on 'datadate' column.
  mutate(month = floor_date(datadate, unit = "month")) %>% 
  # Filter the rows with month of December 2022, with non-zero returns, 
  # and finite values
  filter(month == "2022-12-01") %>% 
  filter(returns != 0) %>% 
  filter(is.finite(returns))

# Omit the NAs
dataD <- na.omit(dataD)
  
# Create a new column to indicate the decile group based on the returns
dataD$percentiles <- cut(dataD$returns,
                         breaks = quantile(dataD$returns,
                                           probs = seq(0, 1, by = 0.1),
                                           type = 5,
                                           na.rm = TRUE),
                         include.lowest = TRUE,
                         labels = paste0(seq(10, 100, by = 10),"%"))

# Check the first few rows
head(dataD)

```

Choose the stock matching the following criteria:
- highest return in the 60% decile group
  
Display the stock in a pretty table with 8 columns showing this information:
- GVKEY, iid, date, company name, trading volume, closing price, return, decile

```{r selectD}
# Create a new data frame 'highest_dataD', to store the filtered rows with 
# highest return in the 60% decile group, and select the specific 8 columns
highest_dataD <- dataD %>% 
  filter(percentiles == "60%") %>% 
  arrange(desc(returns)) %>%
  select(tic, GVKEY, iid, datadate, conm, cshtrd, prccd, returns, percentiles) %>% 
  # Display the first row of' data frame'highest_dataD'
  head(n = 1)

# Display the results in an appropriate formatted table
kable(highest_dataD, caption = "Highest return in the 6th Decile for December 2022")

```

E

Consider only the month of November 2022, and create decile groups.
```{r decilesE}
# Create a new data frame called 'dataE'
dataE <- data_monthly %>% 
  # Add a new variable 'month' and round each date to the beginning of the 
  # respective month.
  mutate(month = floor_date(datadate, unit = "month")) %>% 
  # Filter the rows representing the month of November 2022, with non-zero returns,
  # and only finite values
  filter(month == "2022-11-01") %>% 
  filter(returns != 0) %>% 
  filter(is.finite(returns))

# Omit the NA values in 'dataE'
dataE <- na.omit(dataE)
  
# Add a new variable 'percentiles' that categorizes the 'returns' into percentiles 
# with the specified labels
dataE$percentiles <- cut(dataE$returns,
                         breaks = quantile(dataA$returns,
                                           probs = seq(0, 1, 0.1),
                                           type = 5,
                                           na.rm = TRUE),
                         include.lowest = TRUE,
                         labels = paste0(seq(10, 100, by = 10),"%"))

# Check the first few rows of 'dataE'
head(dataE)
```

Choose the stock matching the following criteria:
- lowest return in the 30% decile group
  
Display the stock in a pretty table with 8 columns showing this information:
- GVKEY, iid, date, company name, trading volume, closing price, return, decile

```{r selectE}
# Create a new data frame called 'lowest_dataE'
lowest_dataE <- dataE %>% 
  # Filter the percentile of 30%
  filter(percentiles == "30%") %>% 
  # Arrange the data frame based on returns 
  arrange(returns) %>%
  # Select only the information in 8 columns
  select(tic,GVKEY, iid, datadate, conm, cshtrd, prccd, returns, percentiles) %>% 
  # Display the first row of data frame 'lowest_dataE'
  head(n = 1)
lowest_dataE

# Create a table to display 'lowest_dataE' in a pretty format.
kable(lowest_dataE, caption = "Lowest return in the 3rd Decile for November 2022")

```

--END--

